Cars Auctions

### 2023-05-09 paper.Rmd
Add descriptions for the rest of the plots

```{r, knitr::opts_chunk$set(echo = F,warning = FALSE, message = FALSE)}
knitr::opts_chunk$set(echo = TRUE, message=F, warning=F)
library(ggplot2)
library(gridExtra)
library(scales)
library(dplyr)

@@

## Heavily damaged cars are more profitable to resell in Armenia.
```{r, echo=F}
combined_data <- left_join(us_cars, am_cars, by = join_by(x$Make == y$Make, 
                                                             x$Model.Group == y$Model,
                                                             x$Year == y$ProductionYear)) %>% na.omit()

@@

```{r, echo=F}

@@

```{r, echo = F}
summarised_data$profit <- summarised_data$Price.AM - summarised_data$Price.Imported
summarised_data <- summarised_data %>%
  filter(profit > 0)
ggplot(summarised_data,aes(x = profit, y = Damage.Description)) + 
  geom_bar(stat = "identity", fill = "red") + ggtitle("Profit of reselling baed on damage type in USD") +
  xlab("Profit in USD") + ylab("Damage type") + scale_x_continuous(labels = comma)
```

@@

## Japanesse cars are cheaper to repair compared to German cars in Armenia.
```{r, echo=F}

@@

german_cars <- us_cars[us_cars$Make %in% german_car_names, ]
japanesse_cars <- us_cars[!(us_cars$Make %in% german_car_names), ]

@@

ggplot(japanesse_cars, aes(x = Repair.cost)) + 
  geom_density(fill = "lightblue") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Distribution of repair costs for Japanesse cars") +
  labs(x = "Repair cost", y = "Density")

@@

# For the US
ggplot(us_cars, aes(x = Color, y = Est..Retail.Value)) + 
  geom_boxplot(fill = "lightblue") + 
  ylim(c(0, 120000)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Prices of the cars by color in US") +
  labs(x = "Color", y = "Price")

# Damage type and the Price
ggplot(us_cars, aes(x = Damage.Description, y = Price)) + 
  geom_violin(fill = "lightblue") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Prices of the cars by damage type in the US") +
  labs(x = "Damage type", y = "Price")

# Milage and the Price
# For US
ggplot(us_cars, aes(x = Odometer, y = Price)) + 
  geom_point(alpha = 0.3, color = "darkblue") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Prices of the cars by milage in the US") +
  labs(x = "Milage", y = "Price")

# For the Armenia
ggplot(am_cars, aes(x = Mileage, y = Price)) + 
  geom_point(alpha = 0.3, color = "darkblue") + 
  xlim(c(0, 200000)) + 
  ylim(c(0, 250000)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Prices of the cars by mileage in Armenia") +
  labs(x = "Milage", y = "Price")

@@

## The popular cars in US and Armenia are quite different
```{r, echo=F}

@@

>>>>

```{r setup, include=FALSE, echo=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE) 
```

```{r}
library(ggplot2)
library(gridExtra)
library(scales)
library(dplyr)

@@

## Heavily damaged cars are more profitable to resell in Armenia.
```{r}
combined_data <- left_join(us_cars, am_cars, by = join_by(x$Make == y$Make, 
                                                             x$Model.Group == y$Model,
                                                             x$Year == y$ProductionYear)) %>% na.omit()

@@

```{r}

@@

```{r}
summarised_data$profit <- summarised_data$Price.AM - summarised_data$Price.Imported
summarised_data <- summarised_data %>%
  filter(profit > 0)
ggplot(summarised_data,aes(x = profit, y = Damage.Description)) + 
  geom_bar(stat = "identity", fill = "red") + ggtitle("Profit of reselling baed on damage type in USD") +
  xlab("Profit in USD") + ylab("Damage type") + scale_x_continuous(labels = comma)
```

@@

## Japanese cars are cheaper to repair compared to German cars in Armenia.
```{r}

@@

german_cars <- us_cars[us_cars$Make %in% german_car_names, ]
Japanese_cars <- us_cars[!(us_cars$Make %in% german_car_names), ]

@@

ggplot(Japanese_cars, aes(x = Repair.cost)) + 
  geom_density(fill = "lightblue") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Distribution of repair costs for Japanese cars") +
  labs(x = "Repair cost", y = "Density")

```
We can see the distribution plots of the repair costs both for Japanese and German cars. For the Japanese cars we can see that most of the observations are within the region of [0, 20000] dollars. Nevertheless, for the German cars the range is much wider reaching up to 35000-40000 dollars for most of the observations. This supports the claim that in general Japanese cars are cheaper to repair compared to German cars in Armenia.

@@

# For the US
ggplot(us_cars, aes(x = Color, y = Est..Retail.Value)) + 
  geom_boxplot(fill = "lightblue") + 
  ylim(c(0, 120000)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Prices of the cars by color in US") +
  labs(x = "Color", y = "Price")
```
We can see the boxplots of the prices for each color both plotted using Armenian and US cars datasets. We can see that the overall picture is quite similar for both countries. However, for Armenia we can notice that Golden color has the highest median compared to US where Burgundy has the highest median. In Armenia, however the price difference between colors and their corresponding price ranges is way more noticeable as such colors as Black and White also have comparably higher prices. Note that we don't take into account the colors which have very few observations.

```{r}
# Damage type and the Price
ggplot(us_cars, aes(x = Damage.Description, y = Price)) + 
  geom_violin(fill = "lightblue") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Prices of the cars by damage type in the US") +
  labs(x = "Damage type", y = "Price")

```
The violin plots above show the distributions of the price of the car by the damage types. As we can see for such damages as All Over, Mechanical and Minor the Price doesn't go higher than 20000 dollars.Yet, for all the other damage types the prices may go up to 58000 dollars. Therefore, we may claim that the damage type has an impact on the price.

```{r}
# Milage and the Price
# For US
ggplot(us_cars, aes(x = Odometer, y = Price)) + 
  geom_point(alpha = 0.3, color = "darkblue") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Prices of the cars by milage in the US") +
  labs(x = "Mileage", y = "Price")

# For the Armenia
ggplot(am_cars, aes(x = Mileage, y = Price)) + 
  geom_point(alpha = 0.3, color = "darkblue") + 
  xlim(c(0, 200000)) + 
  ylim(c(0, 250000)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Prices of the cars by mileage in Armenia") +
  labs(x = "Mileage", y = "Price")

@@

The scatterplots above show the relationship of the Mileage and the price of the cars in Armenia and US. Though we have much less observations for US cars, the overall picture is very much similar. We may notice that we have cars with higher price when the milage is low and as the Milage increases we have less and less cars with higher prices. Therefore, we may state that Mileage does have an impact on the price.

## The popular cars in US and Armenia are quite different
```{r}

@@

```
The barplots above displays the counts of the cars by car models for Armenia and US. For most of the cars which don't have high counts the results are very much alike (Acura, Audi, Mercedes-Benz, Volkswagen). For Toyota, it's nearly the same, since this is the car which appears the most in the auctions both for Armenia and US. Yet, for such car models as BMW, Lexus, Mazda the picture is different for Armenia compared to US, since the numbers are higher for the former. In US, Toyota is followed by Nissan and Honda.
